---
title: "Problem3"
format: html
editor: visual
---

The NFL has two conferences, the NFC and the AFC. Within each conference there are four divisions; East, West, North, and South. Each division has four teams. The Carolina panthers play in the NFC South division. In this problem we will analyze data on NFL results. Use the nflfastR library to download play by play results for the 2023 season. Then do the following:

```{r}
library(nflfastR)
```

• Calculate the total yards by team by game for each team over the course of the **regular** season. (Hint: group by game_date, posteam, and defteam and sum the field yards_gained. Note that under certain conditions the yards_gained or the posteam may be NA so adjust accordingly)

```{r}
regular_season_data <- load_pbp(2023)%>%
  filter(season_type == "REG")
```

```{r}
regular_season_yards <- regular_season_data %>%
  group_by(game_id,game_date, posteam, defteam)%>%
  filter(!is.na(posteam),
         !is.na(defteam)) %>%
  summarize(total_yards_gained=sum(yards_gained, na.rm = TRUE),
            .groups = "drop")
```

• Calculate the minimum, average, median, maximum, and standard deviation of yards per game across the league and report those results.

```{r}
quant_season <- 
  regular_season_yards %>%
  group_by(game_id) %>%
  summarize (Min = min(total_yards_gained, na.rm = TRUE),
             Avg = mean (total_yards_gained, na.rm = TRUE),
             Med = median(total_yards_gained, na.rm = TRUE),
             Max = max (total_yards_gained, na.rm = TRUE),
             StdDev = sd(total_yards_gained, na.rm = TRUE),
             posteam=posteam,
             defteam=defteam,
             .groups="drop")
quant_season
```

• Generate a histogram of the yards per game across the league and record that in the template.

```{r}
regular_season_yards %>%
  ggplot(aes(x=total_yards_gained))+
  geom_histogram()+
  labs(title="Histogram of Yards Per Game Across NFL in 2023 Regular Season",
       x= "Yards Per Game",
       y= "Frequency")
```

• Load the file NFL_Divisions and join that to the Total Yards dataframe.

```{r}
library(readr)
NFL_Divisions <- read_csv("NFL Divisions.csv")
NFL_Divisions
total_yard_teams <- full_join(regular_season_yards, NFL_Divisions, by = join_by("posteam" == "abbr"))
total_yard_teams <- full_join(total_yard_teams, NFL_Divisions, by = join_by("defteam" == "abbr"))

total_yard_teams
```

• Calculate summary statistics (min, average, median, max, std dev) for the total yards per game by division. Record those statistics in the template.

```{r}
quant_teams <- 
  total_yard_teams %>%
  group_by(game_id) %>%
  summarize (Min = min(total_yards_gained, na.rm = TRUE),
             Avg = mean (total_yards_gained, na.rm = TRUE),
             Med = median(total_yards_gained, na.rm = TRUE),
             Max = max (total_yards_gained, na.rm = TRUE),
             StdDev = sd(total_yards_gained, na.rm = TRUE),
             .groups="drop")
quant_teams
```

• Generate a boxplot of total yards per game by division and record that in the template.

• Conduct an ANOVA test to evaluate the null hypothesis that there is no difference in the yard per game across divisions. Record the summary in the template.

• Filter the data to NFC South and calculate summary statistics (min, average, median, max, std dev) for Total Yards by team and report that in the template.

• Generate a box plot of total yards for teams in the NFC South and record that in the template.

• Conduct an ANOVA test to evaluate the null hypothesis that there is no difference in the yard per game across teams in the NFC South. Record the summary in the template.

• Conduct a pairwise Tuckey test to determine the differences for each pair of teams. Report the result in your template.

• Generate a column chart of total yards by game for the Panthers. Paste the graph into the template.

• What can you conclude from this analysis about the Carolina Panthers season?
